---
title: "Exploratory Data Analysis"
output: html_notebook
---

## Setup

This notebook will be used to explore data from the 1850 U.S. Census. This data taken from the [NHGIS](https://www.nhgis.org/).

```{r setup, include=FALSE}
library(tidyverse)
library(skimr)
census <- read_csv("census_1850_data.csv")
codebook <- read_csv("census_1850_codebook.csv")
```

The data frame `census` contains the data. The data frame `codebook` explains more about the column names.

## Summary or skim

We can use `summary()` or `skim()` to get a sense of the data in any column.

```{r}
summary(census$population_total)
skim(census$population_total)
```

This also works on whole data frames, but it is best to do this with a limited selection of columns.

```{r}
census %>% 
  select(starts_with("crops_")) %>%
  skim()
```

## Aggregating by state

We will be better off in some instances getting aggregates by states. We can do this for a single column.

```{r}
census %>% 
  group_by(STATE) %>% 
  summarize(crops_rice = sum(crops_rice, na.rm = TRUE)) %>% 
  arrange(desc(crops_rice))
```

But if we want to do this for a **bunch** of columns, we will be better off with `summarize_at()`.

```{r}
census %>% 
  group_by(STATE) %>% 
  summarize_at(vars(starts_with("crops_")), sum, na.rm = TRUE) 
```

Sometimes it is more illuminating to have a rank than a figure.

```{r}
census %>% 
  group_by(STATE) %>% 
  summarize_at(vars(starts_with("crops_")), sum, na.rm = TRUE) %>% 
  mutate_at(vars(starts_with("crops_")),
            funs(dense_rank(desc(.))))
```

## Single-variate distributions

For a categorical variable we can get a sense of the distribution using a bar plot.

```{r}
ggplot(census, aes(x = STATE)) +
  geom_bar() +
  coord_flip()
```

For a numeric variable we can get the distribution using a histogram. Often you want to zoom in on the x-axis.

```{r}
ggplot(census, aes(x = population_total)) +
  geom_histogram(binwidth = 1000) + 
  coord_cartesian(xlim = c(0, 50e3))
```

Often times you want to adjust for either geography or population. In other words, crops per person or crops per square kilometer.

```{r}
ggplot(census, aes(x = population_total / area_sq_km)) +
  geom_histogram(binwidth = 5) +
  coord_cartesian(xlim = c(0, 200))
```

## Bi-variate correlation

You can compare two numeric variables with `geom_point()` (or `geom_jitter()`, `geom_count()`, or `geom_hex()`).

```{r}
ggplot(census,
       aes(x = population_white_free, 
           y = population_nonwhite_slave)) +
  geom_jitter(shape = 1) 
```

## Exploratory Data Analysis

Explore a variety of variables, using both single-variate distributions and bi-variate correlations.
