---
title: "Judges Example"
---

```{r}
library(tidyverse )
library(igraph)
library(ggraph) 
library(stringr)
library(forcats)
library(lubridate)
library(historydata)
library(tidygraph)
```

Get the data frames from the judges table.

```{r}
data("judges_people")
data("judges_appointments")
judges_people
judges_appointments
```

We are going to remove judge labels from court names, so that the court names are consistent. We are also going to parse the commission date.

```{r}
judges_cleaned_up <- judges_appointments %>% 
  mutate(court_name = court_name %>% 
           str_replace_all("Associate Judge, ", "") %>% 
           str_replace_all("Chief Judge, ", "") %>% 
           str_replace_all(", Chief Judge", "") %>% 
           str_replace_all(", Chief Justice", "")) %>% 
  mutate(commission_date = mdy(commission_date),
         commission_year = year(commission_date)) %>% 
  mutate(president_party = fct_other(president_party,
                                     keep = c("Democratic", "Federalists",
                                              "Jeffersonian Republican",
                                              "Republican", "Whig")))

early_appt <- judges_cleaned_up %>% 
  filter(commission_year <= 1920)
```

Let's create an edge list. The most important thing is that our edges be the first two columns.

```{r}
appt_edges <- early_appt %>% 
  left_join(judges_people, by = "judge_id") %>% 
  select(judge_id, court_name, president_party)
```

```{r}
appt_g <- graph_from_data_frame(appt_edges, directed = FALSE) %>% as_tbl_graph()
appt_g
plot(appt_g, vertex.label = NA, vertex.size = 1)
```

Let's do the plot a better way:

```{r}
ggraph(appt_g, layout = "igraph", algorithm = "nicely") +
  geom_edge_fan() +
  geom_node_point() +
  labs(title = "Judges and Courts") +
  theme_graph()
```
  
We can improve this plot by cleaning up changing the colors, for instance.

```{r}
ggraph(appt_g, layout = "igraph", algorithm = "nicely") +
  geom_edge_fan(aes(color = president_party)) +
  geom_node_point() +
  labs(title = "Judges and Courts") +
  theme_graph()
```

No matter what we do the plot will be tricky because it has both courts and judges in it. So let's project the graph into a single kind of node.

```{r}
type_of_node <- bipartite_mapping(appt_g)
V(appt_g)$type <- type_of_node$type
bi_g <- bipartite_projection(appt_g)
judges_g <- bi_g[[1]]
courts_g <- bi_g[[2]]
```

Let's check in on our courts.

```{r}
courts_g <- courts_g %>% as_tbl_graph()
courts_g
```


We can get the graph data frames back from the `courts_g`, and do a join back to the appointments table to get the data about the courts.

```{r}
courts_info <- judges_appointments %>% 
  select(court_name, court_type) %>% 
  distinct()
courts_g_edges <- igraph::as_data_frame(courts_g, "edges") 
courts_g_nodes <- igraph::as_data_frame(courts_g, "vertices") %>% 
  left_join(courts_info, by = c("name" = "court_name"))
```

Then we can make a new graph. Notice that this time we pass in both an edges list and a node list.

```{r}
courts_g2 <- graph_from_data_frame(courts_g_edges, directed = TRUE, 
                                   vertices = courts_g_nodes) %>% 
  as_tbl_graph()
```

And now we can plot the relationship between courts.

```{r}
ggraph(courts_g2, "igraph", algorithm = "nicely") +
  geom_edge_fan() + 
  geom_node_point(aes(color = court_type))
```

Finally, we can pull out the sub-graph connected to the Supreme Court.

```{r}
induced_subgraph(courts_g2, 14) %>% plot()
```

We can query the graph for various information about its structure without plotting it:

```{r}
articulation_points(courts_g2)
assortativity_degree(courts_g2)
betweenness(courts_g2) %>% sort(decreasing = TRUE) %>% head(5)
closeness(courts_g2) %>% sort(decreasing = TRUE) %>% head(5)
```

